D-Hammer: Efficient Equational Reasoning for Labelled Dirac Notation

### 2025-04-10 05tool_eval.tex
add ref

\subsubsection*{Equivalence checking of circuits}
Quantum circuits are concrete computation models for qubit systems, and they can be described as unitary operators in Dirac notation. 
D-Hammer can check the equivalence of small scaled quantum circuits through their Dirac notation representations. Nevertheless, this is not the intended application. 
Its performance is not comparable with specialized algorithms and tools based on ZX-calculus or decision diagrams, which utilize the specific information of circuit classes to achieve better optimizations. 
An evaluation of some examples is given in~\Cref{fig: circuit eval}, and~\Cref{fig: QC-5} shows one of them. D-Hammer will always time out on larger quantum circuit examples.

\begin{figure}[h]
    \center
    % \scriptsize
    \setlength{\extrarowheight}{2pt}
    \begin{tabular}{c @{\hspace{2em}} c @{\hspace{2em}} c @{\hspace{2em}} c}
        \hline
        example & D-Hammer & ZX-Calculus & Decision Diagrams \\
        \hline
        QC-1 & 0.029 & 0.0 & 0.0 \\
        QC-2 & 0.16 & 0.00039 & 0.0038 \\
        QC-3 & 0.29 & 5.7e-5 & 0.0057 \\
        QC-4 & 0.013 & 4.3e-5 & 0.0017 \\
        QC-5 & 15 & 6.4e-5 & 0.0044 \\
        QC-6 & timeout & 0.00014 & 0.016 \\
        \hline
    \end{tabular}
    \caption{Time consumptions (in seconds) for quantum circuit equivalence checking using different tools.}
    \vspace{-0.2cm}
    \label{fig: circuit eval}
\end{figure}

>>>>

\subsubsection*{Equivalence checking of circuits}
Quantum circuits are concrete computation models for qubit systems, and they can be described as unitary operators in Dirac notation. 
D-Hammer can check the equivalence of small scaled quantum circuits through their Dirac notation representations. Nevertheless, this is not the intended application. 
Its performance is not comparable with specialized algorithms and tools based on ZX-calculus~\cite{9868772} or decision diagrams\cite{10.1145/3394885.3431590}, which utilize the specific information of circuit classes to achieve better optimizations. 
An evaluation of some examples is given in~\Cref{fig: circuit eval}, and~\Cref{fig: QC-5} shows one of them. D-Hammer will always time out on larger quantum circuit examples.

\begin{figure}[h]
    \center
    % \scriptsize
    \setlength{\extrarowheight}{2pt}
    \begin{tabular}{c @{\hspace{2em}} c @{\hspace{2em}} c @{\hspace{2em}} c}
        \hline
        example & D-Hammer & ZX-Calculus~\cite{9868772} & Decision Diagrams (simulation)~\cite{10.1145/3394885.3431590} \\
        \hline
        QC-1 & 0.029 & 0.0 & 0.0 \\
        QC-2 & 0.16 & 0.00039 & 0.0038 \\
        QC-3 & 0.29 & 5.7e-5 & 0.0057 \\
        QC-4 & 0.013 & 4.3e-5 & 0.0017 \\
        QC-5 & 15 & 6.4e-5 & 0.0044 \\
        QC-6 & timeout & 0.00014 & 0.016 \\
        \hline
    \end{tabular}
    \caption{Time consumptions (in seconds) for quantum circuit equivalence checking using different tools.}
    \vspace{-0.2cm}
    \label{fig: circuit eval}
\end{figure}
